### 2012-10-11 thesis/tex/5_design.tex
Initial commit of thesis

\begin{description}
	\item[(a)] \emph{AND} operator: $\varphi \wedge \psi$.
	\item[(b)] \emph{OR} Operator: $\varphi \vee \psi$.
	\item[(c)] \emph{IF} operator: $\varphi \Rightarrow \psi$.\\Instead of an \emph{IMPLIES} operator an \emph{IF} operator is provided. Its meaning seems to be more intuitive for non-experts since it is a common construct in almost every programming language.
	\item[(d)] \emph{NOT} operator: $\neg \varphi$.
	\item[(e)] Proposition: $\rho$.\\Until now there is only the \emph{state proposition} type, which gives evidence about the currently active state. Other types such as numeric or string equations may be added but are less important for our healthcare subject.
	\item[(f)] \emph{ALWAYS} operator: $\Box \varphi$.\\$\varphi$ must be true now and in all following states.
	%\Box \Diamond \medcircle
	\item[(g)] \emph{NEXT} operator: $\medcircle \varphi$.\\$\varphi$ has to be true in the next state.
	\item[(h)] \emph{FUTURE} operator: $\Diamond \varphi$.\\At least one time -- now or in a later state -- $\varphi$ must be true.
	\item[(i)] \emph{UNTIL} operator: $[\varphi \mathcal{U} \psi]$.\\Now and in all following states $\varphi$ must be true until there is a state with $\psi$ being true. In addition eventually there has to be a future state with $\psi$ being true.
\end{description}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.5]{table}
  \caption{The visual language consists of eight unary or binary operator types and one state proposition, each of them having its own color flavor. Logical operators are arranged in a vertical row wereas time relevant statements are aligned horizontally.}
  \label{fig:operators}
\end{figure}

@@

\begin{itemize}
	\item Each operator has a different significant visual appearance; in this case we have used color. This kind of classification enables a faster subconscious identification.
	\item The layout of constraints (nested operators) has two dimensions with a certain meaning. Figure~\ref{fig:directions} demonstrates the ``logical'' vertical read direction as well as the time releveant horizontal line.
	\item Multiple nested operators of the same kind which have no order priorities such as \emph{OR} or \emph{AND} appear unnecessarily complex. These operators can be merged.
\end{itemize}

@@

\begin{itemize}	
	\item All editing is performed only by simple drag and drop. Operators and propositions can be moved and dropped onto other operators. A tool bar provides all operator and proposition types to be instantiated and there is a trash can for deleting existing operators.
	\item In order to simplify the reading of constraints an operator can be hovered over by the mouse pointer. All higher operators become bleached out so that the hovered operator and its children appear highlighted. It's intended to be similar to line coloring when reading a digital text.
	\item Every edit to either the constraint or the underlying state machine results in an automatically (re-)validation of the constraint. The result is immediately displayed to the user.
	\item There is no button for constraint creation / editing except operator and proposition providers, which let the developer create new instances. Having no additional buttons reduces complexity and keeps it simple.
\end{itemize}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.5]{directions}
  \caption{Easy understanding of visual constraints due to intuitive read directions: ``ALWAYS is true: IF state \emph{'Not taken yet'} is active THEN state \emph{'Polling'} can NOT be visited UNTIL state \emph{'Staff notified'} OR state \emph{'Well done!'} is reached.''}
  \label{fig:directions}
\end{figure}

%\begin{figure}[htbp]
  %\centering
  %\includegraphics[scale=0.4]{and_simplify} %width=0.7\linewidth
  %\caption{In order to simplify the appearance of constraints certain operators might be merged. Merged nested \emph{AND} operators (center) are less complex as default ones (left). However nested \emph{AND} operators with mouse over (right) are displayed separated in order to retain full editability.}
  %\label{fig:nested_and}
%\end{figure}

\subsection{Automated constraint generation}
\label{sec:automatedconstraintgeneration}

In order to support the developer in specifying safety constraints, automated generation and suggestion of constraints might be helpful. But it's helpful only if the suggested constraints are relevant and reasonable. This however depends on the domain and use case.
Regarding the important constraints postulated in Sect.~\ref{sec:requirements}, all the relevant states fall into one of the following categories:
%following types of states are of a special interest:

@@

In case of $\alpha = \beta$ -- i.e. when all paths starting from $\alpha$-state eventually return again (cycle) -- even another constraint can be assumed to be true:

@@

All formulas can be translated to the visual formalism presented in section~\ref{sec:visualformalism} and shown to the user. The appropriate LTL formulas for the constraints postulated in section~\ref{sec:requirements} relate to number \ref{eq:second} and \ref{eq:fifth} of the subgraph based formula templates and are as follows:

>>>>

\begin{description}
	\item[(a)] \emph{AND} operator: $\varphi \wedge \psi$.
	\item[(b)] \emph{OR} Operator: $\varphi \vee \psi$.
	\item[(c)] \emph{IF} operator: $\varphi \Rightarrow \psi$.\\Instead of an \emph{IMPLIES} operator an \emph{IF} operator is provided. Its meaning seems to be more intuitive for non-experts since it is a common construct in almost every programming language.
	\item[(d)] \emph{NOT} operator: $\neg \varphi$.
	\item[(e)] Proposition: $\rho$.\\Until now there is only the \emph{state proposition} type, which gives evidence about the currently active state. Other types such as numeric or string equations may be added but are less important for our healthcare subject.
	\item[(f)] \emph{ALWAYS} operator: $\Box \varphi$.\\$\varphi$ must be true now and in all following states.
	%\Box \Diamond \medcircle
	\item[(g)] \emph{NEXT} operator: $\medcircle \varphi$.\\$\varphi$ has to be true in the next state.
	\item[(h)] \emph{FUTURE} operator: $\Diamond \varphi$.\\At least one time - now or in a later state - $\varphi$ must be true.
	\item[(i)] \emph{UNTIL} operator: $[\varphi \mathcal{U} \psi]$.\\Now and in all following states $\varphi$ must be true until there is a state with $\psi$ being true. In addition eventually there has to be a future state with $\psi$ being true.
\end{description}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{table}
  \caption{The visual language consists of eight unary or binary operator types and one state proposition, each of them having its own color flavor. Logical operators are arranged in a vertical row wereas time relevant statements are aligned horizontally.}
  \label{fig:operators}
\end{figure}

@@

\begin{itemize}
	\item Each operator has a different significant visual appearance; in this case we have used color. This kind of classification enables a faster subconscious identification.
	\item The layout of constraints (nested operators) has two dimensions with a certain meaning. Figure~\ref{fig:directions} demonstrates the ``logical'' vertical read direction as well as the time releveant horizontal line.
	\item Multiple nested operators of the same kind which have no order priorities such as \emph{OR} or \emph{AND} appear unnecessarily complex. These operators can be merged (see~figure~\ref{fig:nested_and}).
\end{itemize}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{and_simplify} %width=0.7\linewidth
  \caption{In order to simplify the appearance of constraints certain operators might be merged. Merged nested \emph{AND} operators (center) are less complex as default ones (left). However nested \emph{AND} operators with mouse over (right) are displayed separated in order to retain full editability.}
  \label{fig:nested_and}
\end{figure}

@@

\begin{itemize}	
	\item All editing is performed only by simple drag and drop. Operators and propositions can be moved and dropped onto other operators. A tool bar provides all operator and proposition types to be instantiated and there is a trash can for deleting existing operators.
	\item In order to simplify the reading of constraints an operator can be hovered over by the mouse pointer. All higher operators become bleached out so that the hovered operator and its children appear highlighted. It's intended to be similar to line coloring when reading a digital text.
	\item Every edit to either the constraint or the underlying state machine results in an automatically (re-)validation of the constraint. The result is immediately displayed to the user.
	\item There are different statuses which inform about validation result: For syntactically invalid constraints, an ``incomplete'' sign is displayed. Otherwise, an animated ring indicates that validation is in pro\-gress and will finally result in either a ``valid'' or ``invalid'' sign (see~figure~\ref{fig:results}).
	\item There is no button for constraint creation / editing except operator and proposition providers, which let the developer create new instances. Having no additional buttons reduces complexity and keeps it simple.
\end{itemize}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{directions}
  \caption{Easy understanding of visual constraints due to intuitive read directions: ``ALWAYS is true: IF state \emph{'Not taken yet'} is active THEN state \emph{'Polling'} can NOT be visited UNTIL state \emph{'Staff notified'} OR state \emph{'Well done!'} is reached.''}
  \label{fig:directions}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{results}
  \caption{Status indicators displaying validation results: incomplete, validation in pro\-gress, invalid and valid.}
  \label{fig:results}
\end{figure}

\section{Automated constraint generation}
\label{sec:automatedconstraintgeneration}

As already mentioned before, it might be difficult to find all reasonable safety constraints needed for full safety covery. Hence it also can happen, that a prior software version was correct and over time mistakes crept in because of minor adjustments once in a while.
To help prevent this to happen, we introduce the concept of automated constraint generation for state machines. Once a program reaches a complete condition, constraint suggestions which are valid on the current program can be computed fully automatically. On the one hand they may help the developer identifying reasonable constraints he didn't think about as well as contradictions between program and specification, eg. if a computed constraint doesn't make sense at all. On the other hand all constraints can be revalidated after each change to the software and thus ensure consistency.

In order to support the developer in developing safe programs, automated generation and suggestion of constraints might be helpful. But it's helpful only if the suggested constraints are relevant and reasonable. This however depends on the domain and use case. It has to be mentioned that our approach presented in this work has a special focus on the healtcare use cases and may not be applicable for other domains.
Regarding the important constraints postulated in chapter~\ref{chap:goals}, all the relevant states fall into one of the following categories:

@@

In case of $\alpha = \beta$ - i.e. when all paths starting from $\alpha$-state eventually return again (cycle) - even another constraint can be assumed to be true:

@@

All formulas can be translated to the visual formalism presented in section~\ref{sec:operatorconstraints} and shown to the user. The appropriate LTL formulas for the constraints postulated in chapter~\ref{chap:goals} relate to number \ref{eq:second} and \ref{eq:fifth} of the subgraph based formula templates and are as follows:

### 2012-10-11 thesis/tex/7_prototypeandevaluation.tex
Initial commit of thesis

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\linewidth]{stm.pdf} %width=0.7\linewidth
  \caption{Simplified screen-flow of the medication reminder application.
  % ** COMMENT: THIS ONE IS TOO SMALL! SHOULD MAKE IT BIGGER TO MAKE TEXT READABLE **
  }
  \label{fig:medicationreminder}
\end{figure}

\subsection{Visual constraints}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.4]{sampleconstraint}
  \caption{Visual constraint creation for LTL formula ~\ref{eq:sampleconstraint}.}
  \label{fig:sampleconstraint}
\end{figure}

After each change in the dashboard, the constraint is automatically recompiled and revalidated by using the underlying model checker. For syntactically invalid constraints, an ``incomplete'' sign is displayed in the respective tab. Otherwise, an animated ring indicates that validation is in progress and will finally result in either a ``valid'' or ``invalid'' sign.

%\begin{figure}[htbp]
%  \centering
%  \includegraphics[scale=0.5]{results}
%  \caption{Status indicators displayed during or after validation: incomplete, validation in progress, invalid and valid.}
%  \label{fig:results}
%\end{figure}

@@

\subsection{Constraint generation}
\label{sec:prototypeandevaluation_constraintgeneration}

>>>>

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\linewidth]{stm.pdf} %width=0.7\linewidth
  \caption{Simplified screen-flow of the medication reminder application.}
  \label{fig:medicationreminder}
\end{figure}

\section{Operator constraints}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{sampleconstraint}
  \caption{Visual constraint creation for LTL formula ~\ref{eq:sampleconstraint}.}
  \label{fig:sampleconstraint}
\end{figure}

After each change in the dashboard, the constraint is automatically recompiled and revalidated by using the underlying model checker. For syntactically invalid constraints, an ``incomplete'' sign is displayed in the respective tab. Otherwise, an animated ring indicates that validation is in pro\-gress and will finally result in either a ``valid'' or ``invalid'' sign.

@@

\subsection{Expressiveness}

Ausdruckssärke gleich zu LTL, das heißt, dass lediglich Möglichkeiten nicht ausgedrückt werden können. Zudem gilt eine weitere Einschränkung, dass als bedingungen bisher nur Aussagen der Form ``ist in Zustand x'' verwendet werden können. Da die Validierung der healtbot anwendungen derzeit ausschließlich diesen typ braucht, sind andere Bedingungen wie Stringvergleich oder numerische Gleichheit bisher mit der visuellen Sprache nicht möglich. In Figure~\ref{fig:example_propositions} ist aufgezeigt, wie eine Erweiterung des LTLCreators um weitere Bedingungstypen aussehen könnte.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{example_propositions}
  \caption{Possible proposition extensions: numeric and string equation.}
  \label{fig:example_propositions}
\end{figure}

\section{Automated Constraint generation}
